Deployable Probabilistic Programming

### 2019-04-17 paper/paper.tex
in progress related work

Probabilistic programming is traditionally associated with
design and implemenetation of probabilistic programming
languages. An apparent justification for a new programming
language is that a probabilistic program has a different
semantics~\cite{GHNR14} than a `regular' program of similar
structure. Goodman and Stuhlm\"uller~\cite{GS15} offer a
systematic approach to design and implementation of
probabilistic programming languages based on extension of the
syntax of a (subset of) regular programming language and
transformational compilation to continuation-passing form. Van
de Meent et al~\cite{MPY+18} give a comprehensive overview of
available choices of design and implementation of first-order
and higher-order probabilistic programming languages.
Stan~\cite{Stan17} is built around an imperative probabilistic
programming language with program structure tuned for most
efficient execution of inference on the model.
SlicStan~\cite{GGS19} is built on top of Stan as a
source-to-source compiler, providing the model developer with
a more intuitive language while retaining performance benefits
of Stan.

Integrated probabilistic programming (Figaro~\cite{P09},
Anglican~\cite{TMY+16}, Turing~\cite{GXG18}).

>>>>

Probabilistic programming is traditionally associated with
design and implementation of probabilistic programming
languages. An apparent justification for a new programming
language is that a probabilistic program has a different
semantics~\cite{GHNR14} than a `regular' program of similar
structure. Goodman and Stuhlm\"uller~\cite{GS15} offer a
systematic approach to design and implementation of
probabilistic programming languages based on extension of the
syntax of a (subset of) regular programming language and
transformational compilation to continuation-passing form. Van
de Meent et al~\cite{MPY+18} give a comprehensive overview of
available choices of design and implementation of first-order
and higher-order probabilistic programming languages.
Stan~\cite{Stan17} is built around an imperative probabilistic
programming language with program structure tuned for most
efficient execution of inference on the model.
SlicStan~\cite{GGS19} is built on top of Stan as a
source-to-source compiler, providing the model developer with
a more intuitive language while retaining performance benefits
of Stan. Our work differs from existing approaches in that we
advocate enabling probabilistic programming in a general purpose
programming language by leveraging existing capabilities of the
language, rather than building a new language on top or besides
an existing one.

The need for integration between probabilistic programs and the
surrounding software environment is well understood in the
literature~\cite{TMY+16,Pyro18}. Probabilistic programming
languages are often implemented as embedded domain-specific
languages~\cite{SGG15,GXG18,TMY+16} and include a mechanism of calling
functions from libraries of the host languages. Our work goes a
step further in this direction: since the language of
implementation of probabilistic models coincides with the host
language, any library or user-defined function of the host
language can be directly called from the probabilistic model,
and model methods can be directly called from the host language.
